---
Title: "Pt_modeling"
Author: "Felipe Velásquez"
date: "2024-08-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(rgl)
library(plot3D)

# Random variables as inputs.

n = 2000000         # Number of iterations
set.seed(123)     # Seeds

a = runif(n, 0, 5)                # a = time in Ma
bx = runif(0.8, n)             # bx = Pt Efficiency (%)
b = runif(n, 2, 4)                # b= H2O in parent magma (wt.%)
c = runif(n, 0.2, 1)              # c = H2O in crustal rocks(wt.%)
d = runif(n, 3, 10)              # d = Pressure (kbars)
e = rep(50000, n)                  # e = magma rate (m/Ma)
g = rep(6000, n)                # g = Disk radius (m)
j = runif(n, 50, 100)        # j = Kd Pt by Fleet et al.(1999)_63(17), 2611–2622 
k = runif(n, 0.00096, 0.0058)      # K = Pt total melt (ppm) - metasomatized mantle
                                 

#---------------------------------------------------------------------------------------
# Data generating process:
#---------------------------------------------------------------------------------------

#Crustal Melt Calculations
#--------------------------------------------------------------------------------------

f = a*e                           # f= Magma rate time integrated (m/Ma)

hist(f, breaks = 75)
#---------------------------------------------------------------------------------------
#l= x crustal

l = -0.00000476*(a^{5}) + 0.00010012*(a^{4}) - 0.00087381*(a^{3}) + 0.00390394*(a^{2}) - 0.0087404*(a) + 0.00542933

hist(l, breaks = 75)

#---------------------------------------------------------------------------------------
#m= y crustal

m = 0.00025805*(a^{5}) - 0.00451301*(a^{4}) + 0.03096005*(a^{3}) -
  0.1047359*(a^{2}) + 0.17540876*(a) - 0.08277333

hist(m, breaks = 75)

#---------------------------------------------------------------------------------------
#nn= x crustal
nn = -0.00183672*(a^{5}) + 0.02924904*(a^{4}) - 0.1769496*(a^{3}) +
  0.50100047*(a^{2}) - 0.64552174*(a) + 0.2529533

hist(n, breaks = 75)

#---------------------------------------------------------------------------------------
#o= crustal melt fraction Annen 
o = l*(d^{2}) + m*(d) + nn

hist(o, breaks = 75)

#---------------------------------------------------------------------------------------
#p= crustal Melt fraction Annen logical test
p = 0
for(i in 1:n){
  if(o[i] < 0){
    p[i] = 0
  }
  else{
    p[i] = o[i]
  }
}

hist(p, breaks = 75)

#---------------------------------------------------------------------------------------
#q=crustal Melt fraction Annen logical test2 low
q = 0
for(i in 1:n){
  if((a[i] > 1) & (d[i] > 3)){
    q[i] = p[i]
  }
  else{
    q[i] = NA
  }
}

hist(q, breaks = 75)

#---------------------------------------------------------------------------------------
#r=crustal Melt fraction Annen logical test3
r = 0
for(i in 1:n){
  if((a[i] > 0.25) | (d[i] > 6)){
    r[i] = q[i]
  }
  else{
    r[i] = NA
  }
}

hist(r, breaks = 75)

#---------------------------------------------------------------------------------------
#s=volume of crustal melt fraction rate time integrated (m^3)
s = pi*(g^{2})*(f)*r 

hist(s, breaks = 75)
#---------------------------------------------------------------------------------------
#t= tons of crustal melt fraction time integrated
t = s*(2.6)

hist(t, breaks = 75)


#---------------------------------------------------------------------------------------
# u= crustal Melt percent Annen

u = r*100


#---------------------------------------------------------------------------------------
# Residual Melt Calculations
#---------------------------------------------------------------------------------------
#v= x residual
v = -0.00011309*(a^{6}) + 0.00186017*(a^{5}) - 0.01179089*(a^{4}) +
  0.03594339*(a^{3}) - 0.05316974*(a^{2}) + 0.03207473*(a) + 0.00018778

hist(v, breaks = 75)

#---------------------------------------------------------------------------------------
#w= y residual
w = 0.00120531*(a^{6}) - 0.01970898*(a^{5}) + 0.12379992*(a^{4}) -
  0.37149275*(a^{3}) + 0.53160183*(a^{2}) - 0.28832793*(a) - 0.00343741

hist(w, breaks = 75)

#---------------------------------------------------------------------------------------
# x= z residual
x = -0.00264884*(a^{6}) + 0.04269856*(a^{5}) - 0.26252652*(a^{4}) +
  0.7613096*(a^{3}) - 1.02505321*(a^{2}) + 0.52268727*(a) + 0.01527694

hist(x, breaks = 75)

#---------------------------------------------------------------------------------------
#y= residual Melt fraction Annen
y = v*(d^{2}) + w*(d) + x

hist(y, breaks = 75)
#---------------------------------------------------------------------------------------
#z= residual Melt fraction Annen logical test
z = 0

for(i in 1:n){
  if(y[i] < 0){
    z[i] = 0
  }
  else{
    z[i] = y[i]
  }
}

hist(z, breaks = 75)

#---------------------------------------------------------------------------------------
#aa= residual Melt fraction Annen logical test2
aa = 0
for(i in 1:n){
  if((a[i] > 0.5) & (d[i] > 3)){
    aa[i] = z[i]
  }
}

#--------------------------------------------------------------------------------------
#ab= volume of residual melt fracton time integrated

ab = pi*(g^{2})*(f)*aa

hist(ab, breaks = 75)

#---------------------------------------------------------------------------------------
# ac: tons of residual melt fraction time integrated

ac = 2.6*ab

hist(ac, breaks = 75)

#---------------------------------------------------------------------------------------
#ad= residual Melt percent Annen

ad = aa*100

hist(aa, breaks = 75)

#---------------------------------------------------------------------------------------
#Total Melt Calculations

#---------------------------------------------------------------------------------------
#ae= all_melt_percent_low
ae = u + ad

hist(ae, breaks = 75)

#---------------------------------------------------------------------------------------
#af= Total melt in tons

af = ac + t

#---------------------------------------------------------------------------------------
#ag= Volume km^3 total melt

ag = (af/2.5)/(10^{9})

hist(af, breaks = 75)

#---------------------------------------------------------------------------------------
#H2O Content Calculations in residual melt

#-------------------------------------------------------------------------------------------
#aj= H2O wt.% in residual melt

aj = 100*b*(ad^{-1})

#---------------------------------------------------------------------------------------
#al= r H2O

al =0.00000163*(d^{2})+0.00002024*(d)-0.00000718

hist(al, breaks = 75)

#---------------------------------------------------------------------------------------
#am= s H2O

am = -0.00088884*(d^{2})-0.0110706*(d)+0.00391827

hist(am, breaks = 75)

#---------------------------------------------------------------------------------------
# an: t H2O

an = 0.05908101*(d^{2}) + 1.91102201*(d) + 1.79358317

hist(an, breaks = 75)

#---------------------------------------------------------------------------------------
#ao: residual melt Annen logical test for H2O solubility (F=100-10%)

ao = 0

for(i in 1:n){
  if(is.na(ad[i]) == FALSE){
    if(ad[i] <= 10){
    ao[i] = 10
    }
  else if(ad[i] > 10){
    ao[i] = ad[i]
    }
  }
  else{
    ao[i] = NA
  }
}

hist(ao, breaks = 75)

#----------------------------------------------------------------------------------------
# ak: log test H2O wt% in residual melt
ak = 100*(b)*(ao^{-1})

hist(ak, breaks = 75)

#--------------------------------------------------------------------------------------
# ap: H2O solubility in melt (F =100-10 %)

ap = al*(ao^{2}) + am*ao +  an

hist(ap, breaks = 75)

#---------------------------------------------------------------------------------------------
# aq: H2O wt% in residual melt logical test 100-10% F range

aq = 0

for(i in 1:n){
  if((is.na(aj[i]) == FALSE) & (is.na(ap[i]) == FALSE)){
    if(aj[i] > ap[i]){
      aq[i] = ap[i]
    }
    else if(aj[i] <= ap[i]){
      aq[i] = aj[i]
    }
  }
  else{
    aq[i] = NA
  }
}


#---------------------------------------------------------------------------------------------
# ar: Excess residual H2O

ar = ak - ap

hist(ar, breaks = 75)

#--------------------------------------------------------------------------------------------
# at: Excess H2O residual (tons)

at = (ac*ar)/100

hist(at, breaks = 75)

#--------------------------------------------------------------------------------------------
# au: Log test excess residual H2O tons

au = 0

# =SI(AT3>0,AT3,0)

for(i in 1:n){
  if(is.na(at[i]) == FALSE){
    if(at[i] > 0){
      au[i] = at[i]
    }
    else{
      au[i] = 0
    }
  }
  else{
    au[i] = 0
  }
}

hist(au, breaks = 75)

#-------------------------------------------------------------------------------------------
# as: H2O residual (tons)

as  = (ac-au)*aq/100

#-----------------------------------------------------------------------------------
## H2O Content Calcutation in Crustal Partial Melt

#------------------------------------------------------------------------------
# av: H2O wt% in partial melt

av = 100*(c)*(u^{-1})

#------------------------------------------------------------------------
# aw : H2O solubility (wt.%) with Annen crustal melt % 

aw = (al*(u^{-2})) + (am*u) + an

#------------------------------------------------------------------------
# ax: H2O solubility (wt.%) with Annen crustal melt % logical test

ax = 0

for(i in 1:n){
  if((is.na(av[i]) == FALSE) & (is.na(aw[i]) == FALSE)){
    if(av[i] > aw[i]){
      ax[i] = av[i]
    }
    else if(av[i] <= aw[i]){
      ax[i] = aw[i]
    }
  }
  else{
    aq[i] = NA
  }
}

#------------------------------------------------------------------------
# AY: Excess H2O partial melt

ay = av - aw

hist(ay, breaks = 75)

#------------------------------------------------------------------------

# ba: Excess H2O crustal (tons)

ba = t*ay/100

#-------------------------------------------------------------------------
# BB: log test excess crustal H2O tons

# =SI(BA2>0,BA2,0)

bb = 0

for(i in 1:n){
  if((ba[i] > 0) & (is.na(ba[i])==FALSE)){
    bb[i] = ba[i]
  }
  else{
    bb[i] = 0
  }
}

hist(bb, breaks = 75)

#------------------------------------------------------------------------
#az = H2O crustal (tons)

az = (t-bb)*ax/100

#-------------------------------------------------------------------------

#ah= M res hydrous melt Ton

ah = ac-au+t-bb

#-------------------------------------------------------------------------

#ai=Volume Km^3 hydrous melt

ai = (ah/2.5)/1000000000

#-----------------------------------------------------------------
#-------------------------------------------------------------------------
# bc: H2O total in melt (tons)

bc = as + az


#-------------------------------------------------------------------------
# bd: H2O% in hydrous melt

bd = bc/ah*100


#-------------------------------------------------------------------------
# be: exsolvable H2O tons

be = as + az


#-------------------------------------------------------------------------
# bf: Excess H2O total in melt (tons)

bf = at + ba

hist(bf, breaks = 75)


#-------------------------------------------------------------------------
# bg: Excess total H2O NEW log test

bg = au + bb

hist(bg, breaks = 75)

#-------------------------------------------------------------------------
# bh: M melt - excess H20 tons

bh = ac - au + t + bb

hist(bh, breaks = 75)


#-------------------------------------------------------------------------
# bi: % H2O excess

bi = af - bh

hist(bi, breaks = 75)

#-------------------------------------------------------------------------
# bj: Moles of excess H2O

bj = (bf*1000000)/(2*1.0079 + 15.99)

hist(bj, breaks = 75)

#-------------------------------------------------------------------------



### PLATINUM ENDOWMENT CALCULATIONS

```{r}


# 1 - bk: Pt tot ton

bk = (k*af)/1000000

hist(bk, breaks = 75)

#---------------------------------------------------------------------
# 2 - bl: Pt ppm res melt

bl = bk/((j*af)/1000000)

hist(bl, breaks = 75)


#------------------------------------------------------------------

# 3 - Pt tot tons

bm = (bl*ah)/1000000

bn = bm*0.8

hist(bn, breaks = 75)

summary(bn)

summary(ai)

#------------------------------------------------------------------------

```


```{r}
platinum <- data.frame(cbind(a, k, g, d, j, af, ag, ai, bc, bn))

platinum %>%
  ggplot(aes(x = ai, y = d, colour = a)) + 
  geom_point(aes(colour = a)) + 
  xlab("Volume hydrous melt Km^3") + 
  ylab("Pressure (kbars)") + 
  scale_colour_gradientn(colours = topo.colors(n))

platinum %>%
  ggplot(aes(x=ai)) +
  geom_histogram() +
  xlab("Volume hydrous melt Km^3") + ylab("Frequency")
  
platinum %>%
  ggplot(aes(x = bn, y = d, colour = ai)) +
  geom_point(aes(colour = ai)) + xlab("Pt endowment (tons)") +
  ylab("Pressure (kbars)") +
   scale_colour_gradientn(colours = topo.colors(n))

platinum %>%
  ggplot(aes(x=bn)) +
  geom_histogram() +
  xlab("Platinum Endowment (tons)") + ylab("Frequency")
  
platinum %>%
  ggplot(aes(x = bn, y = ai, colour = a)) +
  geom_point(aes(colour = a)) + xlab("Pt Endowments") +
  ylab("Volume hydrous melt (km3)") +
  scale_colour_gradientn(colours = topo.colors(n))


```

